---
title: "PythonLibrary"
author: "C S 303E UT Austin"
output:
  html_document: 
    theme: darkly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
.watch-out {
  background-color: black;
  border: 2px solid grey;
  font-size: 10px;
  font-weight: bold;
  text-align: left;
  }
```

***

## **Use word search shortcuts like CTRL+G to find specific functions faster**

***

# [**Libraries and packages:**]{style="color:#FF3333;"}

## Libraries

```{r class.source="watch-out"}
# import os : interact with the operating system (e.g. change directory)
# import math : access special math functions such as log(), sin(), sqrt(), pi
# import random : random number generation
# import datetime : clock and calendar functions


# There is a way to use functions from modules(libraries) without calling the module first
# from random import random : allows the use of random() without the random. before it
# from random import * : imports all names(functions) in the module(library)


# libraryName.functionName.__doc__ : returns what the program does (function purpose)
```

***

# [**System functions:**]{style="color:#FF3333;"}

## Standard nifty functions

```{r class.source="watch-out"}
# os.getcwd() : get current working directory - os library
```

***

# [**How data is stored:**]{style="color:#FF0066;"}

## Bytes, memory, file types

```{r class.source="watch-out"}
# All data in the computer is stored as a series of bits (0s and 1s)
# Images can be stored digitally in any of the following formats 
# JPEG: Joint Photographic Experts Group
# PNG: Portable Interchange Format
# GIF: Graphics Interchange Format
# TIFF: Tagged Image File
# PDF: Portable Document Format
# EPS: Encapsulated Postscript

# Memory can be thought of as a big array of bytes (1byte = a sequence of 8 bits)
# Each memory address has an address (o..maximum address) and contents (8 bits)
# A byte is the smallest unit of storage a programmer can address. We say that memory is byte-addressable

# The standard way to represent characters in memory is ASCII 
# The standard ASCII table defines 128 character codes (form 0 to 127)
# A string is represented in memory by a sequence of ASCII character codes (01001010)(J) (01100001)(a)

# Characters or small numbers can be stored in one byte. If data can't be stored in a single byte (e.g., a large number), it must be split across a number of adjacent bytes in memory.
```

***

# [**Characters:**]{style="color:#FF0066;"}

## Operating on characters/value/ordering of characters

```{r class.source="watch-out"}
# lowercase letters have higher ASCII values than uppercase letters (useful to sort)
# the difference between a lowercase and uppercase character is 32

# ord(c) : give the ASCII code for character c; returns a number
# chr(n) : give the character with ASCII code n; returns a character
```

## Special string phrases

```{r class.source="watch-out"}
# Special characters within quotation marks(invalid syntax)
# \b : backspace
# \t : tab
# \n : new line
# \f : formfeed
# \r : carriage return
# \\ : backslash
# \' : single quote
# \" : double quote
```

***

# [**Data Conversion:**]{style="color:#FF0066;"}

## Variable type conversion

```{r class.source="watch-out"}
# int(<number, variable, string>) : Will truncate all that comes after the decimal
# float(<number, variable, string>)
# str(<number, variable>)
# round(<number or variable>) 
# bool(number) if number==0 then False if number >0 then true


# Mixed-type experessions
# Combining two floats result in a float
# Combining two ints results in an int except for (/)
# dividing two ints gives a float 
# combing a float with an int usually yields a float.
```

***

# [**Math Operations:**]{style="color:#9966FF;"}

## Standard operations

```{r class.source="watch-out"}
# + Addition Ex: 34+1=35           : adding two floats results in an approximation ~(1.20000000006)
# - Subtraction Ex: 43.0-0.1=33.9
# * Multiplication Ex: 300*30=9000
# / Float division Ex: 1/2=0.5
# // Integer division Ex: 1//2=0
# ** Exponentiation Ex: 4**0.5=2.0
# pow(a, b) return a^b
# % Remainder division Ex: 20%3=2
# abs(x) Absolute value


# Shorthand syntax
# i+=j same as i=i+j
# i-+j
# i*=j
# i/=j
# i//=j
# i**=j
# i%=j


# [Other operations]
# round(x) : return nearest integer
# round(value, decimal_places) : return the float rounded to b places after the decimal
# max(vector) : return largest argument
# min(vector) : return smallest argument
# random.random() : generates random value between 0 and 1 - random library
# math.sqrt(x) : returns square root of x - math library
# math.pi : returns approximation of pi - math library
# math.floor(x) : return rounded down integer - math library
# math.ceil(x) : return rounded up integer - math library
# math.exp(x) : return e^x - math library
# math.log(x) : return the log of x - math library
# math.sin(math.pi) : returns ~0 (will automatically detect degree or radian maybe...) - math library
# math.degrees(math.pi) : returns 180 as it will convert from radian to degree - math library
# math.radians(180) : returns pi as it will convert from degree to radian - math library
# math.acos(x) : arc sine, useful for calculating angles - math library
# math.asin(x) : arc cosine, useful for calculating angles - math library
# random.radiant(a, b) : return a random integer between a and b inclusively - random library
# random.randrange(a, b) : return a random integer between a and b-1 inclusively - random library
```

***

# [**Logical Expressions:**]{style="color:#9966FF;"}

## Booleans

```{r class.source="watch-out"}
# booleans can be stored in variables and when printed out will print (True/False)
# <, >, <=, >=, ==, != are some boolean expressions
# internally Python uses 0 to represent False and 1 to represent True. do this by using int(boolean)
# and it works the other way around with bool(0) giving False and bool(1/>0) giving True
# In a boolean context 0,"" and None at all represent False and *any other value* stands for True


# Warnings:
# using "==" when comparing floats can be tricky when dealing with floating point error 
# for example (1.1x3 == 3.3) is false because (3.00000003 does not equal 3,3)

# in a boolean context, Python wont always return True or False, just something equivalent
# for example in ("" and 14 : returns ' ') and (0 or "abc" : returns 'abc')


# Logical Operators
# not : logical negation
# and : logical conjunction
# or : logical disunction
# for example (boolean-expression) and not (boolean-expression)
```

## If statements

```{r class.source="watch-out"}
# general form:
# if boolean-expression:
#   statement(s) *must be indented*

# if boolean-expression:
#   true-case-statement(s)
# else:
#   false-case-statement(s)

# if boolean-expression:
#   if boolean-expression:
#     statement(s)

# if boolean-expression1:
#   statement(s)
# elif boolean-expression2:
#   statement(s)
# elif boolean-expression3:
#   statement(s)
# (will do any statement(s) of the first expression which is true)

# if (condition which returns TRUE or FALSE)
#   code to be executed
```

## conditional expressions

```{r class.source="watch-out"}
# variable = value1 if (boolean-expression) else value2 : sets variable to either value1 or value2 based on the expression
```

## Precedence

```{r class.source="watch-out"}
# The following items follow higher precedence the higher on the chart
# +, - (Unary plus, minus)
# **
# not
# *, /, //, %
# +, - (Binary plus, minus)
# <, <=, >, >=
# ==, !=
# and
# or
# Operators on the same line have the same precedence and evaluate left to right
# Use parentheses to override precedence
```

***

# [**Strings and Numeric manipulation:**]{style="color:#9966FF;"}

## formatting fields of strings and numbers

```{r class.source="watch-out"}
# format (value, format-string) : generates a string representation of value formatted as indicated by the format-string
# format(numerical_value, "x.nf") : creates a string representing a float("f) right justified in a field of width 10(aligned right) with 2 decimal digits 
# format(numerical_value, "x.ne") : scientific notation
# format(numerical_value, "x.n%") : percentage
# format(numerical_value, "<x.nf") : ("<") aligned to the left
# format(numerical_value, ">x.nf") : (">") aligned to the right : strings are aligned to left by default
# format(numerical_value, "x.nd") : decimal
# format(numerical_value, "x.nx") : hexadecimal
# format(numerical_value, "x.no") : octal
# format(numerical_value, "x.nb") : binary

# format("string", "20s") : field of 20 as a string

# String + String : will concatenate two strings (StringString) (+= will work the as well) : (cannot add a string and other variable type)
```

***

# [**Input or set values:**]{style="color:#9966FF;"}

## Assigning values to variables

```{r class.source="watch-out"}
# variable=value
# Simultaneous assignments
# m,n = 2,3 #same as m=2 , n=3
# Multiple assignments
# i = j = k = 1 #same as k=1, j=1, i=1
```


## Manual during-program input

```{r class.source="watch-out"}
# variable = input("<prompt string>") : will display "<prompt string>" and wait for user input via keyboard
# variable will then be assigned the user input (AS A STRING)
```

***

# [**Output:**]{style="color:#CC00FF;"}

## Print statements

```{r class.source="watch-out"}
# print("characters/string", : Will output a string no matter the argument variable type
#      sep = "character/s that go inbetween", : "" for no spaces
#      end = "", : print without the implicit new line(stay on same line), can be replaced with any ending phrase
```

***

# [**Date and times:**]{style="color:#3366FF;"}

## Grabing, setting, and displaying time

```{r class.source="watch-out"}
#print(datetime.datetime.now()) : displays current time - datetime library
```

***

# [**Creating programs:**]{style="color:#3366FF;"}

## Defining a program / code group

```{r class.source="watch-out"}
# def Program_Name(): : Defines the program (traditionally named "main")
#   code...
#   return : will end/exit the program
# code must be indented to be considered within the program (tab)
# Program_Name() : Calls the program

# python programFileName.py : can be used within shell to run the program file from shell
# python programFileName.py. ........ to be determined
```

***

# [**Advice on Programming:**]{style="color:#009999;"}

## The process

```{r class.source="watch-out"}
# Requirements >
# Design >
# Implementation >
# Verification >
# Maintenance >
```

***